---
title: "Statistical Inference: Course Project"
output: 
  pdf_document
---
Created by H.Wang on November 19, 2015


## Basic Settings
```{r qplot, fig.width = 4, fig.height=2, message= FALSE}
echo= TRUE  # make scripts visible to others
```

## Description of the Problem

Analyze the ToothGrowth data in the R datasets package, provide data summary and use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 

## Load Data and Perform Basic Exploratory Data Analysis
```{r}
# load the dataset
library(datasets)
data("ToothGrowth")

# convert column dose to factor (same as supp)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Table Check
str(ToothGrowth)
nrow(ToothGrowth)
```

## Provide a basic summary of the data
```{r}
# summary for the dataset
summary(ToothGrowth)

# split of cases between different dose and supp level
table(ToothGrowth$dose, ToothGrowth$supp)

library(ggplot2)
ggplot(data=ToothGrowth, aes(x=dose, y=len, fill=supp)) +
geom_bar(stat="identity",) +  facet_grid(. ~ supp) +  xlab("Dose in Miligrams") + ylab("Tooth Length") + guides(fill=guide_legend(title="Supplement Type"))
```

There is a positive correlation between tooth length and the dose level of VC and OJ.


## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose

Use 95% confidence intervals for two variables and the intercept are as follows:
```{r}
# linear regression for len vs dose and supp
fit <- lm(len ~ dose + supp, data = ToothGrowth)
# confidence intervals
confint(fit)
```

The confidence level is 95% which means 95% of time the simulation results(coefficient estimation) will be in these ranges. All p-values are less than 0.05, rejecting the null hypothesis and suggesting that each variable explains a variability in tooth length. 

```{r}
summary(fit)
```
The intercept is `r fit$coefficients[[1]]`, meaning without supp the average tooth length will be `r fit$coefficients[[1]]` units. The coefficient of dose is `r fit$coefficients[[2]]` . As long as dose is increased by 1 mg (no change in supp) the tooth length would increase `r fit$coefficients[[2]]` units. The last one is for supp VC and it is negative (`r fit$coefficients[[4]]`), which could be interpreted as with delivery a given supp VC (without changing dose) the tooth length will drop `r -fit$coefficients[[4]]` units. We could also conclude that on average, delivery of supp OJ would increase the tooth length by `r -fit$coefficients[[4]]` units due to two category under supp.     

